---
title: "Restaurant_review_HO2"
author: "Bruno Díez Buitrago"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

The goal of this document is to show how to perform different analyses of text documents at a word level, mainly using the tm (text mining) package in R.
This is part of the Intelligent Systems project found on https://github.com/brunaldo07/Hands-On-Project

```{r include=FALSE}
#Including libraries
setwd("F:/informatica/6 Master Data Science/Intelligent Systems/Unit 5/Hands-On-Project/Hands On 1")
library(tm)
library(ggplot2)
library(wordcloud)
library(RWeka)
library(reshape2)
```


# Loading Corpus

We are using a reduced version of the dataset found in https://www.kaggle.com/vigneshwarsofficial/reviews  containing information about restaurant reviews. 
It can be found in the folder "Corpus" of the containing folder "Resources" of the github project.
The corpus containing all instances of the dataset can be found in the folder "reviews" under the "Resources" folder. 

```{r}
source.pos = DirSource("../Resources/reviews", encoding = "UTF-8")
corpus = Corpus(source.pos)
```


# Inspecting the corpus


Let's see the length of the corpus and a the first entries to see the kind of documents it includes

```{r }
length(corpus)
summary(corpus[1:10])

```