---
title: "Restaurant_review_HO2"
author: "Bruno Díez Buitrago"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

The goal of this document is to show how to perform different annotations (word, sentence, part-of-speech, and Penn Treebank parse) over text documents using the openNLP (natural language processing) and the tm (text mining) packages in R.  
This is part of the Intelligent Systems project found on https://github.com/brunaldo07/Hands-On-Project, precisely the second hands on 

```{r include=FALSE}
#Please set your corresponding path to the directory in your terminal
setwd("F:/informatica/6 Master Data Science/Intelligent Systems/Unit 5/Hands-On-Project/Hands On 2")

#Setting up mirror
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)

#Including libraries
# Needed for OutOfMemoryError: Java heap space 
library(rJava)
.jinit(parameters="-Xmx4g")
# If there are more memory problems, invoke gc() after the POS tagging

library(NLP)
install.packages("openNLP")
library(openNLP) 
install.packages("openNLPmodels.en", repos = "http://datacube.wu.ac.at/", type = "source")
library(tm)
library(ggplot2)
install.packages("ggtext")
library(ggtext)
options(java.parameters = "-Xmx8000m")
```


# Auxiliary functions

getAnnotationsFromDocument returns annotations for the text document: word, sentence, part-of-speech, and Penn Treebank parse annotations.
As an alternative, the koRpus package uses TreeTagger for POS tagging.

```{r include = F}
getAnnotationsFromDocument = function(doc){
  x=as.String(doc)
  sent_token_annotator <- Maxent_Sent_Token_Annotator()
  word_token_annotator <- Maxent_Word_Token_Annotator()
  pos_tag_annotator <- Maxent_POS_Tag_Annotator()
  y1 <- NLP::annotate(x, list(sent_token_annotator, word_token_annotator))
  y2 <- NLP::annotate(x, pos_tag_annotator, y1)
  parse_annotator <- Parse_Annotator()
  y3 <- NLP::annotate(x, parse_annotator, y2)
  return(y3)  
} 
```

getAnnotatedMergedDocument returns the text document merged with the annotations.

```{r include = F}
getAnnotatedMergedDocument = function(doc,annotations){
  x=as.String(doc)
  y2w <- subset(annotations, type == "word")
  tags <- sapply(y2w$features, '[[', "POS")
  r1 <- sprintf("%s/%s", x[y2w], tags)
  r2 <- paste(r1, collapse = " ")
  return(r2)  
} 
```

getAnnotatedPlainTextDocument returns the text document along with its annotations in an AnnotatedPlainTextDocument.

```{r include = F}
getAnnotatedPlainTextDocument = function(doc,annotations){
  x=as.String(doc)
  a = AnnotatedPlainTextDocument(x,annotations)
  return(a)  
} 
```

#Loading corpus

We are using a reduced version of the dataset found in https://www.kaggle.com/vigneshwarsofficial/reviews  containing information about restaurant reviews. 
In this hand on, the 1000 reviews set is too big to be processed in a reasonable amount of time so we are using a set of 100 reviews placed on "Resources/short". 

```{r}
source.pos = DirSource("../Resources/short", encoding = "UTF-8")
corpus = Corpus(source.pos)
```


# Inspecting the corpus


Let's see the length of the corpus and a the first entries to see the kind of documents it includes

```{r }
length(corpus)
summary(corpus[1:10])

```
# Annotate corpus
We apply the getAnnotationsFromDocument function to the corpus using lapply function 


```{r}
annotations = lapply(corpus, getAnnotationsFromDocument)
head(annotations[[1]])
tail(annotations[[1]])

```

Let's attach these annotations to que document and store the annotated corpus, as well as storing these annotations inline text
```{r}
corpus.tagged = Map(getAnnotatedPlainTextDocument, corpus, annotations)
corpus.tagged[[1]] 
corpus.taggedText = Map(getAnnotatedMergedDocument, corpus, annotations)
corpus.taggedText[[1]]
```

 
# Exploring annotated documents

We can access and annotatedPlainTextDocument
```{r}
doc = corpus.tagged[[1]] 
doc
```
And then use different functions to see a document, its words, sentences, tagged words and sentences or the parse tree of its sentences
```{r}
as.character(doc)
head(words(doc))
head(sents(doc),1)
head(tagged_words(doc))
head(tagged_sents(doc),1)
head(parsed_sents(doc),1)
```

r