---
title: "Restaurant_review_HO1"
author: "Bruno Díez Buitrago"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

The goal of this document is to show how to perform different analyses of text documents at a word level, mainly using the tm (text mining) package in R.
This is part of the Intelligent Systems project found on https://github.com/brunaldo07/Hands-On-Project

```{r include=FALSE}
#Including libraries
setwd("F:/informatica/6 Master Data Science/Intelligent Systems/Unit 5/Hands-On-Project/Hands On 1")
library(tm)
library(ggplot2)
library(wordcloud)
library(RWeka)
library(reshape2)
```


# Loading Corpus

We are using a reduced version of the dataset found in https://www.kaggle.com/vigneshwarsofficial/reviews  containing information about restaurant reviews
It can be found in the folder "Corpus" of the containing folder "Resources" of the github project

```{r}
source.pos = DirSource("../Resources/Corpus", encoding = "UTF-8")
corpus = Corpus(source.pos)
```


# Inspecting the corpus


Let's see the length of the corpus and a the first entries to see the kind of documents it includes

```{r }
length(corpus)
summary(corpus[1:10])

```
We see it's a group of PlainTextDocument instances
If we inspect them we will see they contain metadata as well

```{r}
inspect(corpus[1])
inspect(corpus[[127]])
```
We can inspect this metadata as well

```{r}
meta(corpus[[127]])

```



# Term Document Matrix
We are going to create a term document matrix and look at its summary
```{r}
tdm = TermDocumentMatrix(corpus)
tdm
```

