---
title: "Restaurant_review_HO1"
author: "Bruno Díez Buitrago"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

The goal of this document is to show how to perform different analyses of text documents at a word level, mainly using the tm (text mining) package in R.  
This is part of the Intelligent Systems project found on https://github.com/brunaldo07/Hands-On-Project , precisely the first hands on

```{r include=FALSE}
#Please set your corresponding path to the directory in your terminal
setwd("F:/informatica/6 Master Data Science/Intelligent Systems/Unit 5/Hands-On-Project/Hands On 1")

#Including libraries
library(tm)
library(ggplot2)
library(wordcloud)
library(RWeka)
library(reshape2)
```


# Loading Corpus

We are using a reduced version of the dataset found in https://www.kaggle.com/vigneshwarsofficial/reviews  containing information about restaurant reviews. 
It can be found in the folder "Corpus" of the containing folder "Resources" of the github project.
The corpus containing all instances of the dataset can be found in the folder "reviews" under the "Resources" folder. 

```{r}
source.pos = DirSource("../Resources/reviews", encoding = "UTF-8")
corpus = Corpus(source.pos)
```


# Inspecting the corpus


Let's see the length of the corpus and a the first entries to see the kind of documents it includes

```{r }
length(corpus)
summary(corpus[1:10])

```
We see it's a group of PlainTextDocument instances
If we inspect them we will see they contain metadata as well

```{r}
inspect(corpus[1])
inspect(corpus[[127]])
```
We can inspect this metadata as well

```{r}
meta(corpus[[127]])

```



# Term Document Matrix
We are going to create a term document matrix and look at its summary
```{r}
tdm = TermDocumentMatrix(corpus)
tdm
```
The high level of sparsity means most of the content of the matrix are zeroes. We can look at an example of this by taking a subset of documents and terms of the TDM.  
In this particular example the subset is :  
- Documents 106 to 110  
- Terms 10 to 30

```{r}
inspect(tdm[106:110,10:30])
```

We can see the number of terms identified in the tdm

```{r}
length(dimnames(tdm)$Terms)

```
Let's see how frequently those terms appear and plot their frequencies ordered 
```{r}
freq=rowSums(as.matrix(tdm))
plot(sort(freq, decreasing = T),col="blue",main="Word frequencies", xlab="Frequency-based rank", ylab = "Frequency")

```
  
We can take a look at the frequencies of the terms

```{r}
head(freq,10)
tail(freq,10)

#most frequent terms
tail(sort(freq),n=15)
```


This number of terms only appear once
```{r}
sum(freq == 1)
```
# Create TDM with transformations and custom stopwords

We can design some custom stopwords in order to skip those terms that are not going to give us much information in the context we are working on (restaurant reviews).
```{r}

review =  corpus[110]
myStopwords = c(stopwords(),"food","restaurant","review")
review = tm_map(corpus[110],removeWords,myStopwords)
review[[1]]$content[1]
```
create new tdm with transformations an custom stop words

```{r}
tdm = TermDocumentMatrix(corpus,
                         control=list(stopwords = myStopwords,
                                      removePunctuation = T, 
                                      removeNumbers = T,
                                      stemming = T))
tdm
```
We can also show the most frequent terms and their frequencies in a bar plot.
```{r}
freq=rowSums(as.matrix(tdm))
high.freq=tail(sort(freq),n=10)
hfp.df=as.data.frame(sort(high.freq))
hfp.df$names <- rownames(hfp.df) 
ggplot(hfp.df, aes(reorder(names,high.freq), high.freq)) +
  geom_bar(stat="identity") + coord_flip() + 
  xlab("Terms") + ylab("Frequency") +
  ggtitle("Term frequencies")
```

# TDM with TF-IDF weights

We take a look at the summary of the tdm resulting from applying TF-IDF weights
```{r}
tdm.tfidf = TermDocumentMatrix(corpus,
                               control = list(weighting = weightTfIdf,
                                              stopwords = myStopwords, 
                                              removePunctuation = T,
                                              removeNumbers = T,
                                              stemming = T))
tdm.tfidf

```
We inspect a subset of the tdm
```{r}
inspect(tdm.tfidf[130:150,100:200])

```

We can also plot the TF-IDF values in order and see the terms with the highest TF-IDF
```{r}
freq=rowSums(as.matrix(tdm.tfidf))

plot(sort(freq, decreasing = T),col="blue",main="Word TF-IDF frequencies", xlab="TF-IDF-based rank", ylab = "TF-IDF")
```
```{r}
tail(sort(freq),n=10)

```
# Association analysis 
Let's see the words most associated with some of the most relean terms
```{r}
asoc.servic = as.data.frame(findAssocs(tdm,"servic", 0.2))
asoc.servic$names <- rownames(asoc.servic)
asoc.servic
```
```{r}
asoc.delici = as.data.frame(findAssocs(tdm,"delici", 0.2))
asoc.delici$names <- rownames(asoc.delici)
asoc.delici
```
Let's see this last result in a bargraph. 
```{r}
ggplot(asoc.delici, aes(reorder(names,delici), delici)) +   
  geom_bar(stat="identity") + coord_flip() + 
  xlab("Terms") + ylab("Correlation") +
  ggtitle("\"delici\" associations")
```
  
Although it is not very useful since all results have the same aspect

# Word cloud
```{r}
pal=brewer.pal(8,"Blues")
pal=pal[-(1:3)]

set.seed(1234)
```

We create an ngram corpus
```{r}
corpus.ngrams = VCorpus(source.pos)

tdm.unigram = TermDocumentMatrix(corpus.ngrams,
                                control=list(stopwords = c(myStopwords,"s","ve"),
                                removePunctuation = T, 
                                removeNumbers = T)) 
```

We obtain the frequency of each term and invoke the wordcloud function

```{r}
word.cloud=wordcloud(words=names(freq), freq=freq,
                     min.freq=25, random.order=F, colors=pal)
```
